% =============================================================================
% JST Paper — Loader Testing Platform
%
% Overleaf: upload the entire docs/paper/ folder (or zip it) as a project.
%           Set this file (main.tex) as "Main document" in Menu -> Settings.
%           The default latexmk compiler handles pdflatex + bibtex automatically.
%
% Local:    pdflatex main && bibtex main && pdflatex main && pdflatex main
%
% Structure:
%   main.tex              -- preamble, title block, abstract, \input sections
%   sections/01..06.tex   -- body sections
%   references.bib        -- bibliography database (BibTeX)
% =============================================================================
\documentclass[10pt,a4paper,twocolumn]{article}

% --- Packages ----------------------------------------------------------------
\usepackage[top=3cm, bottom=2.5cm, left=2.5cm, right=2.5cm, columnsep=0.7cm]{geometry}
\usepackage{times}              % Times New Roman (body)
\usepackage{helvet}             % Arial for abstract/keywords (via \sffamily)
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{listings}
\usepackage{multirow}
\usepackage{array}
\usepackage{tabularx}
\usepackage{cite}
\usepackage{url}

\hypersetup{
    colorlinks=true,
    linkcolor=black,
    citecolor=black,
    urlcolor=blue
}

% --- Caption style -----------------------------------------------------------
\captionsetup{font=small, labelfont=bf}

% --- Listings style ----------------------------------------------------------
\lstset{
    basicstyle=\ttfamily\small,
    breaklines=true,
    frame=single,
    numbers=left,
    numberstyle=\tiny,
    tabsize=2
}

% --- Section formatting (JST style) -----------------------------------------
\usepackage{titlesec}
\titleformat{\section}{\bfseries\normalsize}{\thesection.}{0.5em}{}
\titleformat{\subsection}{\bfseries\normalsize}{\thesubsection.}{0.5em}{}
\titleformat{\subsubsection}{\bfseries\itshape\normalsize}{\thesubsubsection.}{0.5em}{}
\titlespacing*{\section}{0pt}{12pt}{6pt}
\titlespacing*{\subsection}{0pt}{10pt}{4pt}
\titlespacing*{\subsubsection}{0pt}{8pt}{4pt}

% --- Paragraph spacing -------------------------------------------------------
\setlength{\parindent}{1em}
\setlength{\parskip}{0pt}
\setlength{\intextsep}{8pt}

% =============================================================================
\begin{document}

% --- Title + Abstract (kept in main.tex) ------------------------------------
\twocolumn[
\begin{@twocolumnfalse}

\begin{center}
    {\Large\bfseries A Six-Stage Pipeline Model for Systematic Decomposition\\
    and Automated Testing of Shellcode Loaders}
    \vspace{12pt}

    % --- Authors (double-blind: remove before review submission) ---
    {\normalsize
    Author(s)$^{1,*}$\\[4pt]
    $^1$Affiliation(s)\\[4pt]
    $^*$Corresponding author: email@example.com
    }

    \vspace{12pt}
\end{center}

% --- Abstract ----------------------------------------------------------------
\noindent\rule{\textwidth}{0.4pt}
\vspace{4pt}

{\small\sffamily
\noindent\textbf{Abstract.}
Shellcode loaders are central to modern malware, yet knowledge of their internal structure is scattered across blog posts, open-source projects, and conference talks, with no shared framework for systematic comparison.
This paper proposes a six-stage pipeline model that decomposes a shellcode loader into sequential stages---Anti-Analysis, Storage, Allocation, Transformation, Writing, Execution---together with a notation $L_i.T_j$ for identifying techniques within each stage.
Using this model, we implement an automated build platform that generates loader variants through compile-time technique selection and evaluates them in isolated Windows virtual machines.
To isolate the contribution of the loader itself, we keep the payload fixed---a single \texttt{msfvenom} reverse-TCP shellcode---and vary per-stage techniques while measuring detection outcomes against Windows Defender.
As a complementary direction, we outline a stage-aware detection methodology that maps each stage to candidate telemetry sources and rule formats; its rule development and empirical validation are left as future work.
Preliminary results indicate that the Transformation stage dominates static-detection outcomes while Execution techniques influence behavioral detection, supporting the value of stage-level analysis of shellcode loaders.
}

\vspace{4pt}

{\small\sffamily
\noindent\textbf{Keywords:} shellcode loader; pipeline model; malware analysis; automated security testing; detection mapping; technique decomposition
}

\vspace{4pt}
\noindent\rule{\textwidth}{0.4pt}
\vspace{12pt}

\end{@twocolumnfalse}
]

% --- Body sections -----------------------------------------------------------
\input{sections/01_introduction}
\input{sections/02_related_work}
\input{sections/03_methodology}
\input{sections/04_experiments}
\input{sections/05_discussion}
\input{sections/06_conclusion}

% --- Back matter -------------------------------------------------------------
\section*{Acknowledgments}
% TODO: Funding, advisor name.

\section*{Declaration of Competing Interest}
The authors declare no conflicts of interest.

\section*{Author Contributions}
% TODO: CRediT taxonomy.

\section*{AI Disclosure}
% TODO: Disclose AI usage per JST policy.

% --- References --------------------------------------------------------------
\bibliographystyle{ieeetr}
\bibliography{references}

\end{document}
